---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE, message=FALSE, results='hide'}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.path = "man/figures/README-",
  echo = TRUE,
  fig.width = 8,
  fig.height = 6
)

```

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/trending)](https://CRAN.R-project.org/package=trending)
[![Codecov test coverage](https://codecov.io/gh/reconhub/trending/branch/master/graph/badge.svg)](https://codecov.io/gh/reconhub/trending?branch=master)
[![R build status](https://github.com/reconhub/trending/workflows/R-CMD-check/badge.svg)](https://github.com/reconhub/trending/actions)
<!-- badges: end -->

<br>
**<span style="color: red;">Disclaimer</span>**

This package is a work in progress. Please reach out to the
authors before using.

# Trending

*trending* aims to provides a coherent interface to several modelling tools.
Whilst it is useful in an interactive context, it's main focus is to provide
an intuitive interface on which other packages can be developed
(e.g. [*trendbreaker*](https://github.com/reconhub/trendbreaker)).

## Main features

- **Model specification:** Interfaces to common models through intuitive
  functions; `lm_model()`, `glm_model()`, `glm_nb_model`* and `brms_model`**.

- **Model fitting and prediction:** Once specified, models can be fit to data
  and generate confidence and prediction intervals for future data using `fit()`
  and `predict()`.

- **Plotting functionality** A basic plotting method for trending model
  predictions.

\* &nbsp;Requires [MASS](https://CRAN.R-project.org/package=MASS)  
\*\* Requires [brms](https://CRAN.R-project.org/package=brms)

# Installing the package

Once it is released on [CRAN](https://CRAN.R-project.org), you will be able to
install the stable version of the package with:

```{r, eval=FALSE}
install.packages("trending")
```

The development version can be installed from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
if (!require(remotes)) {
  install.packages("remotes")
}
remotes::install_github("reconhub/trending", build_vignettes = TRUE)
```

# Resources

## Vignettes

An overview of *trending* is provided in the included vignette:
* `vignette("Introduction", package = "trending")`

## Websites

The following websites are available:

- The *trending* project on *github*, useful for developers, contributors, and
users wanting to post issues, bug reports and feature requests: <br>
<https://github.com/reconhub/trending>

## Getting help online

Bug reports and feature requests should be posted on *github* using the [*issue* system](https://github.com/reconhub/incidence2/issues). All other questions
should be posted on the **RECON** slack channel see 
<https://www.repidemicsconsortium.org/forum/> for details on how to join.
  