---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE, message=FALSE, results='hide'}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.path = "man/figures/README-",
  echo = TRUE,
  fig.width = 8,
  fig.height = 6
)

```

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/trending)](https://CRAN.R-project.org/package=trending)
[![Codecov test coverage](https://codecov.io/gh/reconhub/trending/branch/master/graph/badge.svg)](https://codecov.io/gh/reconhub/trending?branch=master)
[![R build status](https://github.com/reconhub/trending/workflows/R-CMD-check/badge.svg)](https://github.com/reconhub/trending/actions)
<!-- badges: end -->

<br>
**<span style="color: red;">Disclaimer</span>**

This package is a work in progress. Please reach out to the
authors before using.

# Trending

*trending* aims to provides a coherent interface to several modelling tools.
Whilst it is useful in an interactive context, it's main focus is to provide
an intuitive interface on which other packages can be developed
(e.g. [*trendbreaker*](https://github.com/reconhub/trendbreaker)).

## Main features

- **Model specification:** Interfaces to common models through intuitive
  functions; `lm_model()`, `glm_model()`, `glm_nb_model`* and `brms_model`**.

- **Model fitting and prediction:** Once specified, models can be fit to data
  and generate confidence and prediction intervals for future data using `fit()`
  and `predict()`.

- **Plotting functionality** A basic plotting method for trending model
  predictions.

\* &nbsp;Requires [MASS](https://CRAN.R-project.org/package=MASS)  
\*\* Requires [brms](https://CRAN.R-project.org/package=brms)
  

## Example usage

### An individual model 
```{r, warning=FALSE}
library(outbreaks)  # for data
library(trending)   # for trend fitting
library(dplyr, warn.conflicts = FALSE)  # for data manipulation

# load data
data(covid19_england_nhscalls_2020)

# define a model
model  <- glm_nb_model(count ~ day + weekday)

# select 6 weeks of data (from a period when the prevalence was decreasing)
last_date <- as.Date("2020-05-28")
first_date <- last_date - 8*7
pathways_recent <-
  covid19_england_nhscalls_2020 %>%
  filter(date >= first_date, date <= last_date) %>%
  group_by(date, day, weekday) %>%
  summarise(count = sum(count), .groups = "drop")

# split data for fitting and prediction
dat <-
  pathways_recent %>%
  group_by(date <= first_date + 6*7) %>%
  group_split()

fitting_data <- dat[[2]]
pred_data <- select(dat[[1]], date, day, weekday)

fitted_model <- fit(model, fitting_data)

# confidence and prediction intervals
pred <- predict(fitted_model, pred_data)
glimpse(pred)
plot(pred, "date", fitted_data = fitting_data, fitted_y = "count")
```