---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE, message=FALSE, results='hide'}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.path = "man/figures/README-",
  echo = TRUE,
  fig.width = 8,
  fig.height = 6
)

```

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/trending)](https://CRAN.R-project.org/package=trending)
[![Codecov test coverage](https://codecov.io/gh/reconhub/trending/branch/master/graph/badge.svg)](https://codecov.io/gh/reconhub/trending?branch=master)
[![R build status](https://github.com/reconhub/trending/workflows/R-CMD-check/badge.svg)](https://github.com/reconhub/trending/actions)
<!-- badges: end -->

<br>
**<span style="color: red;">Disclaimer</span>**

This package is a work in progress. Please reach out to the
authors before using.

## An individual model 
```{r, warning=FALSE}
library(trendbreaker)  # for data
library(trending)      # for trends
library(dplyr)         # for data manipulation

# load data
data(nhs_pathways_covid19)

# select last 6 weeks of data
first_date <- max(nhs_pathways_covid19$date, na.rm = TRUE) - 6*7
pathways_recent <- filter(nhs_pathways_covid19, date >= first_date)

# define a linear model
model = lm_model(count ~ day + weekday)

# split data for fitting and prediction
dat <- group_split(pathways_recent, date <= first_date + 5*7)
fitting_data <- dat[[1]]
pred_data <- dat[[2]]

fitted_model <- fit(model, fitting_data)
pred <- predict(fitted_model, pred_data)
glimpse(pred)
```

## Model selection

You can define a number of different regression models using a common interface.
Once defined you can use different strategies to select the
best-fitting/best-predicting model.

As an example we try to predict `hp` of the famous `mtcars` dataset.  Of course,
this is just a toy example. Usually you would use the package to predict counts
data in a time series.

First we define some potential models:
```{r}
stan_cache <- tempfile() # stan compile to c++ and we cache the code
models <- list(
  null = lm_model(hp ~ 1),
  glm_poisson = glm_model(hp ~ 1 + cyl + drat + wt + qsec + am, poisson),
  lm_complex = lm_model(hp ~ 1 + cyl + drat + wt + qsec + am),
  negbin_complex = glm_nb_model(hp ~ 1 + cyl + drat + wt + qsec + am),
  brms_complex = brms_model(
    hp ~ 1 + cyl + drat + wt + qsec + am,
    family = brms::negbinomial(),
    file = stan_cache
  )
)
```

Then we evaluate them using [N-Fold cross validation](https://en.wikipedia.org/wiki/Cross-validation_(statistics)).

```{r, message=FALSE, warning=FALSE}
# we do CV and evaluate three loss function:
# Root-mean-squared error, the huber-loss and mean absolute error.
# The package works with `yardstick` by default.
out <- capture.output( # no log output in readme :)
  auto_select <- select_model(mtcars, models,
    method = evaluate_resampling,
    metrics = list(yardstick::rmse, yardstick::huber_loss, yardstick::mae)
  )
)
auto_select$leaderboard
```
