---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*trending* aims to provides a coherent interface to several modelling tools.
Whilst it is useful in an interactive context, it's main focus is to provide
an intuitive interface on which other packages can be developed
(e.g. [*trendbreaker*](https://github.com/reconhub/trendbreaker)).

# Main features

- **Model specification:** Interfaces to common models through intuitive
  functions; `lm_model()`, `glm_model()`, `glm_nb_model` and `brms_model`*.

- **Model fitting and prediction:** Once specified, models can be fit to data
  and generate confidence and prediction intervals for future data using `fit()`
  and `predict()`.

\* &nbsp; Requires [brms](https://CRAN.R-project.org/package=brms)

# Example usage

## An individual model 
```{r}
library(outbreaks)  # for data
library(trending)   # for trend fitting
library(dplyr, warn.conflicts = FALSE)  # for data manipulation

# load data
data(covid19_england_nhscalls_2020)

# define a model
model  <- glm_nb_model(count ~ day + weekday)

# select 6 weeks of data (from a period when the prevalence was decreasing)
last_date <- as.Date("2020-05-28")
first_date <- last_date - 8*7
pathways_recent <-
  covid19_england_nhscalls_2020 %>%
  filter(date >= first_date, date <= last_date) %>%
  group_by(date, day, weekday) %>%
  summarise(count = sum(count), .groups = "drop")

# split data for fitting and prediction
dat <-
  pathways_recent %>%
  group_by(date <= first_date + 6*7) %>%
  group_split()

fitting_data <- dat[[2]]
pred_data <- select(dat[[1]], date, day, weekday)

fitted_model <- fit(model, fitting_data)

# default
fitted_model %>% 
  predict(pred_data) %>%
  glimpse()

# without prediction intervals
fitted_model %>% 
  predict(pred_data, add_pi = FALSE) %>% 
  glimpse()

# without uncertainty
fitted_model %>% 
  predict(pred_data, uncertainty = FALSE) %>% 
  glimpse()
```

## Multiple models
Trending makes use of [purrr::safely()] to handle lists of models.  In this
situation the fitted output wil be of class [`trending_model_fit_list`] which
is a list with components result and error. If an error occurred, error is an 
error object and result has a NULL value (otherwise). If the model does not
error then result is the corresponding model output and the error entry will be
NULL.

```{r}
models  <- list(
  simple = lm_model(count ~ day),
  glm_poisson = glm_model(count ~ day, family = "poisson"),
  glm_negbin = glm_nb_model(count ~ day + weekday),
  will_error = glm_nb_model(count ~ day + nonexistant)
)

models %>%
  fit(fitting_data) %>%
  glimpse()
```

Similarly [`trending_model_fit_list`] objects can then be used with [predict]
```{r}
models %>%
  fit(fitting_data) %>% 
  predict(pred_data) %>% 
  glimpse()
```


