---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 7,
  fig.height = 5
)
```

*trending* aims to provides a coherent interface to several modelling tools.
Whilst it is useful in an interactive context, it's main focus is to provide
an intuitive interface on which other packages can be developed
(e.g. [*trendbreaker*](https://github.com/reconhub/trendbreaker)).

# Main features

- **Model specification:** Interfaces to common models through intuitive
  functions; `lm_model()`, `glm_model()`, `glm_nb_model` and `brms_model`*.

- **Model fitting and prediction:** Once specified, models can be fit to data
  and generate confidence and prediction intervals for future data using `fit()`
  and `predict()`.

- **Error and warning catching:** Both `fit` and `predict` methods (for
  `trending_model` and `trending_fit` objects) catch all warnings and return
  the output and these captured values in a single-row tibble. These results
  can then be handled with additional generics `is_ok` and `ok` to determine
  whether the fitting / prediction functionality has succeeded and return
  successful results respectively.

\* &nbsp; Requires [brms](https://CRAN.R-project.org/package=brms)

# Example usage

## Setup
```{r message=FALSE}
library(outbreaks)  # for data
library(trending)   # for trend fitting
library(dplyr)      # for data manipulation

# load data
data(covid19_england_nhscalls_2020)

# select 6 weeks of data (from a period when the prevalence was decreasing)
last_date <- as.Date("2020-05-28")
first_date <- last_date - 8*7
pathways_recent <-
  covid19_england_nhscalls_2020 %>%
  filter(date >= first_date, date <= last_date) %>%
  group_by(date, day, weekday) %>%
  summarise(count = sum(count), .groups = "drop")

# split data for fitting and prediction
dat <-
  pathways_recent %>%
  group_by(date <= first_date + 6*7) %>%
  group_split()

fitting_data <- dat[[2]]
pred_data <- select(dat[[1]], date, day, weekday)
```

## A succesful model fit
```{r}
(model  <- glm.nb_model(count ~ day + weekday))
(fitted_model <- fit(model, fitting_data))
is_ok(fitted_model)
ok(fitted_model)

# default
fitted_model %>% 
  predict(pred_data) %>%
  ok()

# without prediction intervals
fitted_model %>% 
  predict(pred_data, add_pi = FALSE) %>% 
  ok()

# without uncertainty
fitted_model %>% 
  predict(pred_data, uncertainty = FALSE) %>% 
  ok()

# non-bootstraped (parametric) prediction intervals
fitted_model %>% 
  predict(pred_data, simulate_pi = FALSE) %>% 
  ok()
```

## Example of failed model fitting
```{r}
(model2  <- glm.nb_model(count ~ day + nonexistent))
(fitted_model2 <- fit(model2, fitting_data))
is_ok(fitted_model2)
ok(fitted_model2)
```
