---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*trending* aims to provides a coherent interface to several modelling tools.
Whilst it is useful in an interactive context, it's main focus is to provide
an intuitive interface on which other packages can be developed
(e.g. [*trendbreaker*](https://github.com/reconhub/trendbreaker)).

# Main features

- **Model specification:** Interfaces to common models through intuitive
  functions; `lm_model()`, `glm_model()`, `glm_nb_model` and `brms_model`*.

- **Model fitting and prediction:** Once specified, models can be fit to data
  and generate confidence and prediction intervals for future data using `fit()`
  and `predict()`.

\* &nbsp; Requires [brms](https://CRAN.R-project.org/package=brms)

# Example usage

## An individual model 
```{r}
library(outbreaks)  # for data
library(trending)   # for trend fitting
library(dplyr, warn.conflicts = FALSE)  # for data manipulation

# load data
data(covid19_england_nhscalls_2020)

# define a model
model  <- glm_nb_model(count ~ day + weekday)

# select 6 weeks of data (from a period when the prevalence was decreasing)
last_date <- as.Date("2020-05-28")
first_date <- last_date - 8*7
pathways_recent <-
  covid19_england_nhscalls_2020 %>%
  filter(date >= first_date, date <= last_date) %>%
  group_by(date, day, weekday) %>%
  summarise(count = sum(count), .groups = "drop")

# split data for fitting and prediction
dat <-
  pathways_recent %>%
  group_by(date <= first_date + 6*7) %>%
  group_split()

fitting_data <- dat[[2]]
pred_data <- select(dat[[1]], date, day, weekday)

fitted_model <- fit(model, fitting_data)

# default
fitted_model %>% 
  predict(pred_data) %>%
  glimpse()

# without prediction intervals
fitted_model %>% 
  predict(pred_data, add_pi = FALSE) %>% 
  glimpse()

# without uncertainty
fitted_model %>% 
  predict(pred_data, uncertainty = FALSE) %>% 
  glimpse()

# non-bootstraped (parametric) prediction intervals
fitted_model %>% 
  predict(pred_data, simulate_pi = FALSE) %>% 
  glimpse()
```

## Multiple models
When working with a list of models Trending captures both errors and warnings
so the user can choose how to proceed.  In this situation the fitted output will
be of class `trending_model_fit_list` which is a list with named lists for
each model containing components result, warnings and errors. If an error
occurs, then the error object is captured and the result has a NULL value.
If the model does not error then result is the corresponding model output and
the error entry will be NULL.  Similarly, the warning component will be a
vector of all warnings that occurred.

```{r}
models  <- list(
  simple = lm_model(count ~ day),
  glm_poisson = glm_model(count ~ day, family = "poisson"),
  glm_negbin = glm_nb_model(count ~ day + weekday),
  will_error = glm_nb_model(count ~ day + nonexistant)
)

res <- models %>%
  fit(fitting_data)

res

res %>% glimpse()

```

`trending_model_fit_list` objects can then be used with `predict()` which
similarly captures warnings and errors and returns the model predictions in a
column called **output**.
```{r}
res <- models %>%
  fit(fitting_data) %>% 
  predict(pred_data)

res

res %>% glimpse()
```


